Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-02-14 3dpas_intro.tex
checkpoint.

From \S\ref{sec:distTypes} and \figurename~\ref{fig:figures_application}, we
find that applications with distributed data vary in the amount or volume of
data distributed, the degree of distribution (scattered over one site or many
sites), and the dynamic aspects associated with distribution, e.g., latency and
relative scheduling flexibility (and thus tolerance or constraints on data
transfer/aggregation).  Applications processing the largest volumes of data
(O(PB)) typically run on specialized data cyberinfrastructure (e.g., the ATLAS
LHC Grid). Applications such as Fusion and NGS that run on smaller data sets
mainly utilize HPC environments. On the other end of the spectrum are highly
dynamic applications that process sensor data and make near-realtime decisions
based on streaming data, e.g., the power grid and sensor network applications.

@@

Note that this term is scale-invariant, i.e., a value of 1 could
correspond to very small or very large values of both numerator and
denominator, and that it does not capture the time scales over
which data changes. However, we believe it captures most of the
non-temporal {\it dynamic} properties.

@@

A survey of data-intensive distributed applications on production
cyberinfrastructure leads us to posit that there are three primary
macroscopic architectures~\cite{pilot_clouds} and approaches for distributed
data-intensive applications: (i) localize all data into a ``cloud''
(or a large analytical engine) -- the paradigm adopted by several
genome projects and the cancer atlas, (ii) decompose and distribute
data to an appropriate number of computing/analytical engines as
available -- the paradigm employed by particle physics for the
discovery of the Higgs Boson (similar to the application in
\S\ref{WLCGSteve}), and (iii) a hybrid of the above two paradigms,
wherein data is decomposed and committed to several infrastructure,
which in turn could be a combination of either of the first two
paradigms, similar to what is done in the Power grids application
(\S\ref{power}). %\jhanote{improve description....replace hybrid by bottom-up? }

It is obvious that the first architecture can be used for large
volumes of data, but surely there are limitations to the scalability
or validity of this model. How does this limitation change,
qualitatively or quantitatively, if dynamic data and/or applications
are considered? This begets the question: how fundamental an issue is
{\it dynamic}?

The high-energy physics (HEP) community has self-organized towards the
second architectural approach.  Although distributed computing
infrastructures representing this approach, such as OSG and EGI, have
been successful for this community and application, broad uptake
across other application types, modes, and data volumes has not taken
place. What lessons does the HEP experience teach us for the other
communities, such as the bioinformatics?  Can an infrastructure
conforming to a particular architecture and developed for a specific
application domain be generalized to another science domain, using the
same underlying macro architecture but by providing a different set of
services and specialized data cyberinfrastructure? Or is the
architecture specific to a {\it fixed} range of compute-data
characteristics? If so, what are these characteristics?

@@

\appendix
\section*{Appendix: Methodology}

>>>>

From \S\ref{sec:distTypes} and \figurename~\ref{fig:figures_application}, we
find that applications with distributed data vary in the amount or volume of
data distributed, the degree of distribution (scattered over one site or many
sites), and the dynamic aspects associated with distribution, e.g., latency and
relative scheduling flexibility (and thus tolerance or constraints on data
transfer/aggregation).  
% Applications processing the largest volumes of data typically run on
% specialized data cyberinfrastructure (e.g., the ATLAS LHC Grid). Applications
% such as Fusion and NGS that run on smaller data sets mainly utilize HPC
% environments.

Some applications use specialized data cyberinfrastructure, while some run
utilize existing HPC environments.  On the other end of the spectrum are highly
dynamic applications that process sensor data and make near-realtime decisions
based on streaming data, e.g., the power grid and sensor network applications.

@@

Note that this term is scale-invariant, i.e., a value of 1 could
correspond to very small or very large values of both numerator and
denominator, and that it does not capture the time scales over
which data changes.

% However, we believe it captures most of the non-temporal {\it dynamic}
% properties.

@@

An investigation of data-intensive distributed applications on production
cyberinfrastructure, leads us to believe that there are three primary
macroscopic architectures and approaches for distributed data-intensive
applications: (i) localize all data into a ``cloud'' (or a large analytical
engine) -- the paradigm adopted by several genome projects and the cancer atlas,
(ii) decompose and distribute data to an appropriate number of
computing/analytical engines as available -- the paradigm employed by particle
physics for the discovery of the Higgs Boson (similar to the application in
\S\ref{WLCGSteve}), and (iii) a hybrid of the above two paradigms, wherein data
is decomposed and committed to several infrastructure, which in turn could be a
combination of either of the first two paradigms, similar to what is done in the
Power grids application
(\S\ref{power}). %\jhanote{improve description....replace hybrid by bottom-up? }

It is obvious that the first architecture can be used for large volumes of data,
but there are self-evident limitations to the scalability or validity of this
model. How does this limitation change, qualitatively or quantitatively, if
dynamic data and/or applications are considered? This begets the question: how
fundamental an issue is {\it dynamic}?

The high-energy physics (HEP) community has self-organized towards the
second architectural approach.  Although distributed computing
infrastructures representing this approach, such as OSG and EGI, have
been successful for this community and application, broad uptake
across other application types, modes, and data volumes has not taken
place. What lessons does the HEP experience teach us for the other
communities, such as bioinformatics?  Can an infrastructure
conforming to a particular architecture and developed for a specific
application domain be generalized to another science domain, using the
same underlying macro architecture but by providing a different set of
services and specialized data cyberinfrastructure? Or is the
architecture specific to a {\it fixed} range of compute-data
characteristics? If so, what are these characteristics?

@@

\newpage
\appendix
\section*{Appendix: Methodology}
